---
title: "Lab 3 Homework"
author: "Jason Nguyen"
date: "Winter 2020"
output:
  html_document: 
    keep_md: yes
    theme: spacelab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run.  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data
For the homework, we will use data about vertebrate home range sizes. The data are in the class folder, but the reference is below.  

**Database of vertebrate home range sizes.**  
Reference: Tamburello N, Cote IM, Dulvy NK (2015) Energy and the scaling of animal space use. The American Naturalist 186(2):196-211. http://dx.doi.org/10.1086/682070.  
Data: http://datadryad.org/resource/doi:10.5061/dryad.q5j65/1  

1. Load the data into a new object called `homerange`.  
```{r}
homerange <- readr::read_csv("data/Tamburelloetal_HomeRangeDatabase.csv")
```

2. Use `spec()` to see the full details of the columns.  
```{r}
spec(homerange)
```

3. Explore the data. Show the dimensions, column names, classes for each variable, and a statistical summary. Keep these as separate code chunks.  
```{r}
dim(homerange)
```

```{r}
colnames(homerange)
```

```{r}
lapply(homerange, class)
#nice, I didn't know this one! You can also use-
#glimpse(homerange)
```

```{r}
summary(homerange)
```


4. Are there NA's in your data? Show the code that you would use to verify this, please.  

```{r}
anyNA(homerange)
```

5. Change the class of the variables `taxon` and `order` to factors and display their levels.  
```{r}
homerange$taxon <- as.factor(homerange$taxon)
class(homerange$taxon)
```

```{r}
levels(homerange$taxon)
```

```{r}
homerange$order <- as.factor(homerange$order)
class(homerange$order)
```

```{r}
levels(homerange$order)
```


6. Make a new dataframe `deer` that is limited to the mean mass, log10 mass, family, genus, and species of deer in the database. The family for deer is cervidae. Arrange the data in descending order by log10 mass. Which is the largest deer?  
```{r}
deer <- homerange %>% 
  filter(family=="cervidae") %>% 
  select(mean.mass.g, log10.mass, family, genus, species) %>% 
  arrange(desc(log10.mass))
deer
```
  The largest deer is Alces alces.

7. As measured by the data, which snake species has the smallest homerange? Show all of your work, please. Look this species up online and tell me about it!  
```{r}
snake <- homerange %>% 
  filter(taxon == "snakes") %>% 
  select(common.name, mean.hra.m2) %>% 
  arrange(mean.hra.m2)
snake
```
Namaqua dwarf adder has the smallest homerange. "Namaqua dwarf adder is the smallest adder in the world... It occurs in coastal dunes in Namaqualand and southern Namibia where it feeds on lizards and Rain frogs... The venom of this snake is mildly cytotoxic causing pain and swelling." [Source](https://www.africansnakebiteinstitute.com/snake/namaqua-dwarf-adder/)

8. You suspect that homerange and mass are correlated in birds. We need a ratio that facilitates exploration of this prediction. First, build a new dataframe called `hra_ratio` that is limited to genus, species, mean.mass.g, log10.mass, log10.hra. Arrange it in ascending order by mean mass in grams.  
```{r}
hra_ratio <- homerange %>% 
  filter(taxon == "birds") %>% 
  select(genus, species, mean.mass.g, log10.mass, log10.hra) %>% 
  arrange(mean.mass.g)
hra_ratio
```


9. Replace the existing `hra_ratio` dataframe with a new dataframe that adds a column calculating the ratio of log 10 hra to log 10 mass. Call it `hra.mass.ratio`. Arrange it in descending order by mean mass in grams.  
```{r}
hra.mass.ratio <- hra_ratio %>%
  mutate(hra.mass.ratio=log10.hra/log10.mass) %>% 
  arrange(desc(mean.mass.g))
hra.mass.ratio
```


10. What is the lowest mass for birds with a `hra.mass.ratio` greater than or equal to 4.0?
```{r}
hra.mass.ratio%>% 
  filter(hra.mass.ratio >= 4.0) %>% 
  arrange(mean.mass.g)
```
The lowest mass for a bird with a `hra.mass.ratio` greater than or equal to 4.0 is 5.15g.

11. Do a search online; what is the common name of the bird from #10 above? Place a link in your markdown file that takes us to a webpage with information on its biology.  

  Regulus regulus is the bird with the lowest mass of the birds that have a hra mass ratio greater than or equal to 4. The common name for this bird is "Goldcrest." For more information, click [here](https://en.wikipedia.org/wiki/Goldcrest)

12. What is the `hra.mass.ratio` for an ostrich? Show your work, please.  
```{r}
homerange %>% 
  filter(common.name == "ostrich") %>% 
  mutate(hra.mass.ratio=log10.hra/log10.mass) %>% 
  select(common.name, log10.hra, log10.mass, hra.mass.ratio)
```
  The hra.mass.ratio for an ostrich is 1.602565.
  
#Remember you calculated mass ratio for all lines in hra.mass.ratio (though this is a little confusing since it is also a varible name, perhaps just calling that object hra_ratio2 might be better) So you could also do-

```{r}
#hra.mass.ratio %>% 
  #filter(genus=="struthio" & species=="camelus")
```


## Push your final code to GitHub!
Please be sure that you have check the `keep md` file in the knit preferences.  